Artificial Ecosystem Selection Update

### 2020-10-22 Presentation/presentation.Rmd
update presentation for lab meeting. update heritability calc to include neutral line.

- Give the rationale/motivation for this project, what I've learned so far, and what I could learn from future data.
- Hear  your questions and comments that might guide future steps.
- Talk about what "heritability" means in a community context.

@@

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
knitr::include_graphics('draught-stability.png')
```

@@

1. There's a clear unit of biological organization - the species. 
2. It's obvious what a "plant" is.
3. Communities are relatively simple - (1-25 species).

@@

  1. Without a consistent species concept, it is not obvious at what level of
     biological organization to evaluate this relationship. (the ASV? the
     genus? the family?)
  2. It is not always clear which members of a community are members of a
     particular functional group. Further, in a metabolic network, it's not
     always obvious which functional group is limiting the end product.
  3. Microbial communities are immensely diverse (100s-1000s of ASVs) and
     without solutions to the
     previous two problems, sifting through that diversity is challenging.

@@

- Generate lots of communities with different combinations of taxa.
- Put them into a common environment to minimize environmental variation.
- Test which combinations generate the greatest function.
- Reshuffle the best communities and test again.
- Similar to dilution to extinction or enrichment culturing.

## Method

To do this, I generated 24 ecosystems.

@@

There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.
<style>
.container{
    display: flex;
}
.col{
    flex: 1;
}
</style>

@@

@@

h2_per_gen_sum  <-
  h2_per_gen %>% 
  summarize(h2_med = median(h2), h2_mad = mad(h2))
```

```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
ggplot(h2_per_gen, aes(x = passage, y = h2)) +
  geom_point() +
  geom_pointrange(aes(x = 2.5, y = h2_med, 
                                  ymax = h2_med + h2_mad,
                                  ymin = h2_med - h2_mad),
  data = h2_per_gen_sum) +
  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                h^2, ")"))) +
  theme_classic()

@@

Each small point is the $h^2$ for that passage and its offspring. The
larger central point is the median $h^2$ (+/- median absolute difference)
across all four passages. $h^2$ =  `r myround(h2_per_gen_sum$h2_med, 2)` $\pm$ `r myround(h2_per_gen_sum$h2_mad, 2)`.

# Conclusions and interpetation

```{r, fig.width=4, fig.height=3, out.width=6e2*.8, out.height=4.5e2*.8}

## `r myround(h2_per_gen_sum$h2_med, 2)` of the variation in methane oxidation rate is attributable to the additive effect of taxa passed from the parents

The remaining `r 1 - as.numeric(myround(h2_per_gen_sum$h2_med, 2))` is due to
environmental variation not controlled for experimentally or with the neutral
treatment, interactions among taxa, ...?

Caveat: this doesn't account for cor

```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
ggplot(h2_per_gen, aes(x = passage, y = h2)) +
  geom_point() +
  geom_pointrange(aes(x = 2.5, y = h2_med, 
                                  ymax = h2_med + h2_mad,
                                  ymin = h2_med - h2_mad),
  data = h2_per_gen_sum) +
  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                h^2, ")"))) +
  theme_classic()

## Next: Which taxa changed as a response to selection?

>>>>

- Practice talking about why I'm doing what I'm doing and what we can learn from it.
- Hear your questions and comments that might guide future steps.
- Talk about what "heritability" means in a community context.

@@

```{r, fig.width=4, fig.height=3, out.width=6e2*.8, out.height=4.5e2*0.8}
knitr::include_graphics('draught-stability.png')
```

@@

1. There's a clear unit of biological organization - the species. 
2. It's obvious what a "plant" is.
3. Communities are relatively simple (1-25 species).

@@

  1. No consistent species concept.
  2. What's a plant?
  3. Diversity can be 100s to 10,000s.

@@

1. Generate lots of communities with different combinations of taxa.
2. Put them into a common environment to minimize environmental variation.
3. Test which combinations generate the greatest function.
4. Reshuffle the best communities and test again.
5. Investigate which taxa respond to selection.

## Design

@@

<style>
.container{
    display: flex;
}
.col{
    flex: 1;
}
</style>

@@

There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.

@@

h2_per_gen_sum <-
  h2_per_gen %>% 
  group_by(treat) %>% 
  summarize(h2_med = median(h2), h2_mad = mad(h2), .groups = 'drop')
```

## Half of the variation in methane oxidation rate is attributable to the additive effect of genes passed from the parents

Positive: `r as.numeric(myround(h2_per_gen_sum$h2_med[2], 2))` 
$\pm$
`r as.numeric(myround(h2_per_gen_sum$h2_mad[2], 2))`

Neutral: `r as.numeric(myround(h2_per_gen_sum$h2_med[1], 2))` 
$\pm$
`r as.numeric(myround(h2_per_gen_sum$h2_mad[1], 2))`

The remainder is due to
environmental variation not controlled for experimentally or with the neutral
treatment, interactions among taxa, ...?

Caveat: $h^2$ may be affected by correlated traits also under selection.

## Heritability for the positive and neutral lines

<style>
.container{
    display: flex;
}
.col{
    flex: 1;
}
</style>

<div class="container">
<div class="col">
```{r, warning=FALSE, fig.width=4, fig.height=3, out.width=6e2*0.8, out.height=4.5e2*0.8}
ggplot(h2_per_gen, aes(x = passage, y = h2, color = treat)) +
  geom_point() +
  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                h^2, ")"))) +
  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
                     values = c('gray60', 'darkorange2')) +
  theme_classic()

@@

<div class="col">

```{r, warning=FALSE, fig.width=4, fig.height=3, out.width=6e2*0.8, out.height=4.5e2*0.8}
ggplot(h2_per_gen, aes(x = passage, y = h2, color = treat)) +
  geom_point() +
  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                h^2, ")"))) +
  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
                     values = c('gray60', 'darkorange2')) +
  theme_classic() +
  ylim(-0.683, 1.56)

# Which taxa respond to selection? Are correlated with flux?

## In progress:

- Sequencing *pmoA* to characterize the diversity of methane oxidizers.
- Sequencing 16S to look broadly at bacterial/archaeal taxa.
